Scaffolder - Software for Microbial Genome Scaffolding.

### 2011-01-31 text/src/article.tex
Spell check

The Scaffolder software allows manual or computational joining of mutliple
discontiguous nucleotide sequences together into a contiguous super sequence.
Simple plain-text configuration files are used to select the nucleotide
sequences in the scaffold and how they should be joined together. This is
called the ``scaffold file''. \pb

@@

The original raw nucleotide sequence for each assembled contig or insert
sequence is maintained in a fasta file. The fasta file of contigs is a common
output for assemblers and as such should be readily available after assembly
without additional processing. The choice of the nucleotide sequence to use in
the scaffold is specified using the first space-delimited word from the fasta
header of the sequence. By maintaining the nucleotide sequences in a separate
file this preserves the original raw sequence in the fasta file. Furthermore
this separates the actual nucleotide sequence from the determination of how it
is selected and editted in the final genome sequence. \pb

@@

The genome scaffold is wriiten in YAML format. YAML is a data format designed
to be both machine readable and easy to manually edit. An example of this
format illustrating all the features of scaffolder is shown in Figure 1. \pb

@@

The first line of the file begins with three dashes. This is required to
signify the start of a YAML document. Lines 2 begins with a dash character `-'
which denotes the start of an entry in the sequence scaffold, in this case the
first entry. Line 3 is indented by two spaces which therefore means this line
contains an entry type. The ``sequence'' tag is used indicate a sequence is
inserted at this region in the scaffold. The following line contain is
indented by two spaces and indicates that this an attribute of the sequence
entry. In this case the ``source'' keyword is used to indicate used which
sequence should be inserted into the scaffold. This sequence is indentified by
the first word of the fasta header of the selected sequence. This indicates
a simple sequence region added to the scaffold. \pb

\subsubsection*{Unresolved seqeunce region} %{{{3

The following line has no indentation and begins with dash indicating the next
entry in the scaffold. The following entry is indentified by the ``unresolved''
tag. This identifies a region of unknown nucleotide sequence but is included
because the approximate size of the region is known. The following line defines
that the size of this unknown region in the genome scaffold should be 20
nucleotides. This therefore results in the region being filled with 20 `N'
characters in draft sequence. \pb

@@

Inserts are listed similarly to sequence entries where reverse, start and stop
attributes can be used to trim and reverse the sequence. The `source' tag is
used to indentify which fasta sequence entry should be used as the insert.  The
insert has two additional tags ``open'' and ``close'' which determine where the
insert is added to the hosting sequence. These coordinates are inclusive
replaceing the coordinate nucleotides. The open and close positions need not be
the same size as the insert size. Any difference in size to the replaced
nucleotides will result in the length on the encoding sequence being adjusted.
\pb

@@

\paragraph{Sub-sequence reverse complemented: } If specified with the reverse keyworkd the sequence is reverse complemented.
\pb

@@

Scaffolder relies on both the ruby programming language and the RubyGems
package management system. These are commonly installed on recent versions of
Mac OSX and linux distrubtions. The installation of these can be tested on the
command line by typing the following commands. \pb

@@

The main aims of scaffolder are two fold. The first is usuability and the second is reproducibility.

@@

Scaffolder is designed to be as simple to use as possible. Assuming the
existence of Ruby and RubyGems, scaffolder can be installed with a single
command. Scaffolder uses a minimal and compact syntax to described the
scaffold. This syntax is simple to construct and edit whilst being succint and
readable in describing how the scaffold was constructed. Scaffolder provides
extensive documentation both for the command line tools and for the Scaffolder
Ruby API.

@@

Constructing a genome sequence using a plain text scaffold file means the same
sequence is always output for the same file. The sequence produced from
manually joining separate contigs in a text editor can not be reliably
reproduced. Constructing the scaffold in a plain text file means that the
sequence is reproducible. Furthermore the source of each region in final
sequence can be identified. The scaffold file provides a readable record of
how the sequence was constructed. The scaffold file also means the scaffold is
easer to edit once already constructed. Much more difficult to do when working
on the sequence alone. The YAML text format is also much easier to compare
differences using standard diff commands. This makes scaffold files much more
simple to store in version control and to see the stages in the construction
of the genome sequence.

@@

Scaffolder follows the Unix philospophy ``Do one thing and do it well''.
Scaffolder focuses on creating a simple to use and install tool for generating
a genome scaffold. The YAML data format for creating a scaffold file is
standardised and easily manipulated programmatically. This thereby follows
a second Unix tenet: ``If your data structures are good enough, the algorithm
to manipulate them should be trivial.''

>>>>

The Scaffolder software allows manual or computational joining of multiple
discontiguous nucleotide sequences together into a contiguous super sequence.
Simple plain-text configuration files are used to select the nucleotide
sequences in the scaffold and how they should be joined together. This is
called the ``scaffold file''. \pb

@@

The original raw nucleotide sequence for each assembled contig or insert
sequence is maintained in a fasta file. The fasta file of contigs is a common
output for assemblers and as such should be readily available after assembly
without additional processing. The choice of the nucleotide sequence to use in
the scaffold is specified using the first space-delimited word from the fasta
header of the sequence. By maintaining the nucleotide sequences in a separate
file this preserves the original raw sequence in the fasta file. Furthermore
this separates the actual nucleotide sequence from the determination of how it
is selected and edited in the final genome sequence. \pb

@@

The genome scaffold is written in YAML format. YAML is a data format designed
to be both machine readable and easy to manually edit. An example of this
format illustrating all the features of scaffolder is shown in Figure 1. \pb

@@

The first line of the file begins with three dashes. This is required to
signify the start of a YAML document. Lines 2 begins with a dash character `-'
which denotes the start of an entry in the sequence scaffold, in this case the
first entry. Line 3 is indented by two spaces which therefore means this line
contains an entry type. The ``sequence'' tag is used indicate a sequence is
inserted at this region in the scaffold. The following line contain is
indented by two spaces and indicates that this an attribute of the sequence
entry. In this case the ``source'' keyword is used to indicate used which
sequence should be inserted into the scaffold. This sequence is identified by
the first word of the fasta header of the selected sequence. This indicates
a simple sequence region added to the scaffold. \pb

\subsubsection*{Unresolved sequence region} %{{{3

The following line has no indentation and begins with dash indicating the next
entry in the scaffold. The following entry is identified by the ``unresolved''
tag. This identifies a region of unknown nucleotide sequence but is included
because the approximate size of the region is known. The following line defines
that the size of this unknown region in the genome scaffold should be 20
nucleotides. This therefore results in the region being filled with 20 `N'
characters in draft sequence. \pb

@@

Inserts are listed similarly to sequence entries where reverse, start and stop
attributes can be used to trim and reverse the sequence. The `source' tag is
used to identify which fasta sequence entry should be used as the insert.  The
insert has two additional tags ``open'' and ``close'' which determine where the
insert is added to the hosting sequence. These coordinates are inclusive
replacing the coordinate nucleotides. The open and close positions need not be
the same size as the insert size. Any difference in size to the replaced
nucleotides will result in the length on the encoding sequence being adjusted.
\pb

@@

\paragraph{Sub-sequence reverse complemented: } If specified with the reverse keyword the sequence is reverse complemented.
\pb

@@

Scaffolder relies on both the ruby programming language and the RubyGems
package management system. These are commonly installed on recent versions of
Mac OSX and Linux distributions. The installation of these can be tested on the
command line by typing the following commands. \pb

@@

The main aims of scaffolder are two fold. The first is usability and the second is reproducibility.

@@

Scaffolder is designed to be as simple to use as possible. Assuming the
existence of Ruby and RubyGems, scaffolder can be installed with a single
command. Scaffolder uses a minimal and compact syntax to described the
scaffold. This syntax is simple to construct and edit whilst being succinct and
readable in describing how the scaffold was constructed. Scaffolder provides
extensive documentation both for the command line tools and for the Scaffolder
Ruby API.

@@

Constructing a genome sequence using a plain text scaffold file means the same
sequence is always output for the same file. The sequence produced from
manually joining separate contigs in a text editor can not be reliably
reproduced. Constructing the scaffold in a plain text file means that the
sequence is reproducible. Furthermore the source of each region in final
sequence can be identified. The scaffold file provides a readable record of
how the sequence was constructed. The scaffold file also means the scaffold is
easier to edit once already constructed. Much more difficult to do when working
on the sequence alone. The YAML text format is also much easier to compare
differences using standard diff commands. This makes scaffold files much more
simple to store in version control and to see the stages in the construction
of the genome sequence.

@@

Scaffolder follows the Unix philosophy ``Do one thing and do it well''.
Scaffolder focuses on creating a simple to use and install tool for generating
a genome scaffold. The YAML data format for creating a scaffold file is
standardised and easily manipulated programmatically. This thereby follows
a second Unix tenet: ``If your data structures are good enough, the algorithm
to manipulate them should be trivial.''
